Fine-scale predictions of the presence of predatory fish using a machine-learning framework

### 2025-03-13 manuscript.qmd
worked on discussion and adding timer for runs

## model interpretability
We used permutational importance to compare feature importance across different models

@@

# Discussion

@@

## method may be species specific
- hard to collect data for species that appear sporadically in large numbers
- in effective for species that are highly mobile

## limitations for "cell" size when collecting data
- sampling area size should reflect the behavior of a fish
- if data transects are too large, then environmental heterogeneity may be overly "averaged"; i.e., each transect may be too similar
- similarly, larger areas are more likely to contain fish, so the model may simply predict fish in every "cell"
- areas that are too small may inflate the number of absences in the data, so the model may simply predict that all cells are empty.

@@

>>>>

## model interpretability
We used permutational importance to compare feature importance across different models [@fig-var-imp].

@@

# Discussion
 - importance of predation on management outcomes
 - invasive speices
 - traditional statistics models can identify key environmental variables but tend not to be general enough for predictions
 - If environments can be adequately characterized then ML can do both
 - this is still early days for applications of such models in htese subjects

 - develpped flexible pipeline to benchmark models
 - say what i found
 - summarize how pipeline is arranged
 - transition to how data collection needs to be standardized/thought out

@@

## method may be species specific
- hard to collect data for species that appear sporadically in large numbers
- in effective for species that are highly mobile
<!-- 
## limitations for "cell" size when collecting data

- sampling area size should reflect the behavior of a fish
- if data transects are too large, then environmental heterogeneity may be overly "averaged"; i.e., each transect may be too similar
- similarly, larger areas are more likely to contain fish, so the model may simply predict fish in every "cell"
- areas that are too small may inflate the number of absences in the data, so the model may simply predict that all cells are empty. -->

@@

Another consideration is the size of each transect. Transects that are too large may have environmental parameters that differ little from the average of the whole study site. Additionally, larger transects are more likely to contain fish, so the model training process may simply converge on models that predict fish at every location. Conversely, using transects that are too small may result in an overwhelming number of absences simply because fish move even within their own territories, resulting in models that predict absences at every location. Therefore, transects should be small enough to contain a heterogenous mix of environmental parameters, while being large enough to be comparable to a fish's territorial size. [mention dataset that failed?] Such modeling approaches may be inappropriate or need additional considerations for species that are highly mobile or seasonally sporadic, such as striped bass in the Sacramento River [ref].
